Application of Equal Local Levels to Improve Q-Q Plot Testing Bands with \proglang{R} Package \pkg{qqconf}

### 2021-07-03 article/manuscript.Rnw
made updates to api

qq_conf_plot(
  obs = pvals_scenario_2,
  distribution = qunif,
  method = "ell",
  alpha = .05,
  samples_func = lines,
  samples_col = "red"
)

@@

qq_conf_plot(
  obs = pvals_scenario_2,
  distribution = qunif,
  method = "ell",
  alpha = .05,
  samples_func = lines,
  samples_col = "red",
  log10 = TRUE
)

@@

\subsection{Testing for Normality of Regression Residuals}
When performing an ordinary least squares regression (OLS), it is common to assume that the error terms are idependentally drawn from a normal distribution. Specifically, suppose that we have $n$ observations $\{\mathbf{x_{i}}, y_{i}\}_{i=1}^{n}$, where each $y_{i}$ is a scalar and each $\mathbf{x_{i}}$ is a $1xp$ vector of regressors. Then, we model each $y_{i}$ as
\begin{equation*}
  y_{i} = \mathbf{x_{i}}^{T}\boldsymbol\beta + \epsilon_{i}
\end{equation*}
where $\mathbf{\boldsymbol\beta}$ is a $1xp$ vector of regression parameters to be estimated and we assume
\begin{equation}
  \epsilon_{1}, ..., \epsilon_{n} \distras{iid} N(0, \sigma^{2}).
\end{equation}
Once $\boldsymbol\beta$ and $\sigma^{2}$ are estimated (typically via maximum liklihood), for each $y_{i}$ we define the residual as 
\begin{equation*}
  r_{i} = \mathbf{x_{i}}^{T}\mathbf{\hat{\boldsymbol\beta}} - y_{i},
\end{equation*}
where $\hat{\boldsymbol\beta}$ is our estimate of $\boldsymbol\beta$. Then, we would like a Q-Q plot of the residuals with a Normal reference distribution to aid in testing assumption (4) above. Strictly speaking, even if assumption (4) is correct, the residuals are not independent because OLS leads to an estimation of $\hat{\boldsymbol\beta}$ such that $\sum_{i=1}^{n}r_{i} = 0$. However, accounting for the resulting negative correlation between the residuals when creating testing bands would be very difficult, and with a sufficiently large sample size this correlation becomes negligible. Thus, without an a priori reason to believe that the errors may deviate from normality at any particular point in the distribution, it makes sense to use ELL bands in this case. This is very easy to do with \pkg{qqconf}, as we show below.
\newline
\newline
First, we generate data to perform a regression. Here, we will generate each $\epsilon_{i}$ independently from a $t(2.5)$ distribution. 
%
<<generate_datas>>=
set.seed(100)
n <- 100
x <- runif(n)
eta <- rt(n, df = 2.5)
y <- x + eta
@
%
Then, we fit a regression with the simulated data
%
<<fit_regression>>=
reg <- lm(y ~ x)
@
%
With \pkg{base}-\proglang{R} functionality, we can can create a qqplot using the function \code{qqnorm} and \code{qqline}. 
%
%
<<base_qqplot, fig=TRUE>>=
qqnorm(reg$residuals)
qqline(reg$residuals)
@
%
%
\newline
Clearly there's some indication of deviation in this plot, but it can be hard to tell how significant the deviation is without confidence bounds. With our function \code{qq_conf_plot}, it is much easier to be able to judge the significance of the deviation.
%
%
<<qqconf_qqplot, fig=TRUE>>=
library(qqconf)
qq_conf_plot(
  obs = reg$residuals,
  samples_func = lines,
  samples_col = "red"
)
@
%
%
\newline
Here, we can clearly see that in the right tail of the residuals goes beyond the normal testing bounds, giving strong evidence that the errors were not generated from a normal distribution.

>>>>

qq_conf_plot(
  obs = pvals_scenario_2,
  distribution = qunif,
  method = "ell",
  alpha = .05,
  points_params = list(col="red", type="l")
)

@@

qq_conf_plot(
  obs = pvals_scenario_2,
  distribution = qunif,
  method = "ell",
  alpha = .05,
  points_params = list(col="red", type="l"),
  log10 = TRUE
)

@@

\subsection{Testing for Normality of Regression Residuals}
When performing an ordinary least squares regression (OLS), it is common to assume that the error terms are idependentally drawn from a normal distribution. Specifically, suppose that we have $n$ observations $\{\mathbf{x_{i}}, y_{i}\}_{i=1}^{n}$, where each $y_{i}$ is a scalar and each $\mathbf{x_{i}}$ is a $1xp$ vector of regressors. Then, we model each $y_{i}$ as
\begin{equation*}
  y_{i} = \mathbf{x_{i}}^{T}\boldsymbol\beta + \epsilon_{i}
\end{equation*}
where $\mathbf{\boldsymbol\beta}$ is a $1xp$ vector of regression parameters to be estimated and we assume
\begin{equation}
  \epsilon_{1}, ..., \epsilon_{n} \distras{iid} N(0, \sigma^{2}).
\end{equation}
Once $\boldsymbol\beta$ and $\sigma^{2}$ are estimated (typically via maximum liklihood), for each $y_{i}$ we define the residual as 
\begin{equation*}
  r_{i} = \mathbf{x_{i}}^{T}\mathbf{\hat{\boldsymbol\beta}} - y_{i},
\end{equation*}
where $\hat{\boldsymbol\beta}$ is our estimate of $\boldsymbol\beta$. Then, we would like a Q-Q plot of the residuals with a Normal reference distribution to aid in testing assumption (4) above. Strictly speaking, even if assumption (4) is correct, the residuals are not independent because OLS leads to an estimation of $\hat{\boldsymbol\beta}$ such that $\sum_{i=1}^{n}r_{i} = 0$. However, accounting for the resulting negative correlation between the residuals when creating testing bands would be very difficult, and with a sufficiently large sample size this correlation becomes negligible. Thus, without an a priori reason to believe that the errors may deviate from normality at any particular point in the distribution, it makes sense to use ELL bands in this case. This is very easy to do with \pkg{qqconf}, as we show below.
\newline
\newline
First, we generate data to perform a regression. Here, we will generate each $\epsilon_{i}$ independently from a $t(2.5)$ distribution. 
%
<<generate_datas>>=
set.seed(100)
n <- 100
x <- runif(n)
eta <- rt(n, df = 2.5)
y <- x + eta
@
%
Then, we fit a regression with the simulated data
%
<<fit_regression>>=
reg <- lm(y ~ x)
@
%
With \pkg{base}-\proglang{R} functionality, we can can create a qqplot using the function \code{qqnorm} and \code{qqline}. 
%
%
<<base_qqplot, fig=TRUE>>=
qqnorm(reg$residuals)
qqline(reg$residuals)
@
%
%
\newline
Clearly there's some indication of deviation in this plot, but it can be hard to tell how significant the deviation is without confidence bounds. With our function \code{qq_conf_plot}, it is much easier to be able to judge the significance of the deviation.
%
%
<<qqconf_qqplot, fig=TRUE>>=
library(qqconf)
qq_conf_plot(
  obs = reg$residuals,
  points_params = list(col="red")
)
@
%
%
\newline
Here, we can clearly see that in the right tail of the residuals goes beyond the normal testing bounds, giving strong evidence that the errors were not generated from a normal distribution.

### 2021-07-03 article/manuscript.tex
made updates to api

\begin{Schunk}
\begin{Sinput}
R> library(qqconf)
R> n <- 1000
R> bounds <- get_bounds_two_sided(alpha = .05, n = 1000)
R> # Plot 3: Q-Q plot of P-values from Chi-Squared Test
R> # for Independence in a 2 x 2 Table
R> 
R> 
R> qq_conf_plot(
+    obs = pvals_scenario_2,
+    distribution = qunif,
+    method = "ell",
+    alpha = .05,
+    samples_func = lines,
+    samples_col = "red"
+  )
R> # Add the Q-Q plot for scenario 1.
R> lines(c(1:n) / (n + 1), sort(pvals_scenario_1), col = "blue")
R> legend(
+    "topleft", 
+    legend = c("s=20","s=200"), 
+    col = c("red","blue"), 
+    lty = 1
+  )
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> # Plot 4: Q-Q plot, on the -log10 Scale, of P-values 
R> # from Chi-Squared Test for Independence in a 2 x 2 Table
R> n <- 1000
R> qq_conf_plot(
+    obs = pvals_scenario_2,
+    distribution = qunif,
+    method = "ell",
+    alpha = .05,
+    samples_func = lines,
+    samples_col = "red",
+    log10 = TRUE
+  )
R> # Add the Q-Q plot for scenario 1.
R> lines(-log10(c(1:n) / (n + 1)), -log10(sort(pvals_scenario_1)), col = "blue")
R> legend(
+    "topleft",
+    legend = c("s=20","s=200"), 
+    col = c("red","blue"), 
+    lty = 1
+  )
\end{Sinput}
\end{Schunk}

@@

\subsection{Testing for Normality of Regression Residuals}
When performing an ordinary least squares regression (OLS), it is common to assume that the error terms are idependentally drawn from a normal distribution. Specifically, suppose that we have $n$ observations $\{\mathbf{x_{i}}, y_{i}\}_{i=1}^{n}$, where each $y_{i}$ is a scalar and each $\mathbf{x_{i}}$ is a $1xp$ vector of regressors. Then, we model each $y_{i}$ as
\begin{equation*}
  y_{i} = \mathbf{x_{i}}^{T}\boldsymbol\beta + \epsilon_{i}
\end{equation*}
where $\mathbf{\boldsymbol\beta}$ is a $1xp$ vector of regression parameters to be estimated and we assume
\begin{equation}
  \epsilon_{1}, ..., \epsilon_{n} \distras{iid} N(0, \sigma^{2}).
\end{equation}
Once $\boldsymbol\beta$ and $\sigma^{2}$ are estimated (typically via maximum liklihood), for each $y_{i}$ we define the residual as 
\begin{equation*}
  r_{i} = \mathbf{x_{i}}^{T}\mathbf{\hat{\boldsymbol\beta}} - y_{i},
\end{equation*}
where $\hat{\boldsymbol\beta}$ is our estimate of $\boldsymbol\beta$. Then, we would like a Q-Q plot of the residuals with a Normal reference distribution to aid in testing assumption (4) above. Strictly speaking, even if assumption (4) is correct, the residuals are not independent because OLS leads to an estimation of $\hat{\boldsymbol\beta}$ such that $\sum_{i=1}^{n}r_{i} = 0$. However, accounting for the resulting negative correlation between the residuals when creating testing bands would be very difficult, and with a sufficiently large sample size this correlation becomes negligible. Thus, without an a priori reason to believe that the errors may deviate from normality at any particular point in the distribution, it makes sense to use ELL bands in this case. This is very easy to do with \pkg{qqconf}, as we show below.
\newline
\newline
First, we generate data to perform a regression. Here, we will generate each $\epsilon_{i}$ independently from a $t(2.5)$ distribution. 
%
\begin{Schunk}
\begin{Sinput}
R> set.seed(100)
R> n <- 100
R> x <- runif(n)
R> eta <- rt(n, df = 2.5)
R> y <- x + eta
\end{Sinput}
\end{Schunk}
%
Then, we fit a regression with the simulated data
%
\begin{Schunk}
\begin{Sinput}
R> reg <- lm(y ~ x)
\end{Sinput}
\end{Schunk}
%
With \pkg{base}-\proglang{R} functionality, we can can create a qqplot using the function \code{qqnorm} and \code{qqline}. 
%
%
\begin{Schunk}
\begin{Sinput}
R> qqnorm(reg$residuals)
R> qqline(reg$residuals)
\end{Sinput}
\end{Schunk}
\includegraphics{manuscript-base_qqplot}
%
%
\newline
Clearly there's some indication of deviation in this plot, but it can be hard to tell how significant the deviation is without confidence bounds. With our function \code{qq_conf_plot}, it is much easier to be able to judge the significance of the deviation.
%
%
\begin{Schunk}
\begin{Sinput}
R> library(qqconf)
R> qq_conf_plot(
+    obs = reg$residuals,
+    samples_func = lines,
+    samples_col = "red"
+  )
\end{Sinput}
\end{Schunk}
\includegraphics{manuscript-qqconf_qqplot}
%
%
\newline
Here, we can clearly see that in the right tail of the residuals goes beyond the normal testing bounds, giving strong evidence that the errors were not generated from a normal distribution.

>>>>

\begin{Schunk}
\begin{Sinput}
R> library(qqconf)
R> n <- 1000
R> bounds <- get_bounds_two_sided(alpha = .05, n = 1000)
R> # Plot 3: Q-Q plot of P-values from Chi-Squared Test
R> # for Independence in a 2 x 2 Table
R> 
R> 
R> qq_conf_plot(
+    obs = pvals_scenario_2,
+    distribution = qunif,
+    method = "ell",
+    alpha = .05,
+    points_params = list(col="red", type="l")
+  )
R> # Add the Q-Q plot for scenario 1.
R> lines(c(1:n) / (n + 1), sort(pvals_scenario_1), col = "blue")
R> legend(
+    "topleft", 
+    legend = c("s=20","s=200"), 
+    col = c("red","blue"), 
+    lty = 1
+  )
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> # Plot 4: Q-Q plot, on the -log10 Scale, of P-values 
R> # from Chi-Squared Test for Independence in a 2 x 2 Table
R> n <- 1000
R> qq_conf_plot(
+    obs = pvals_scenario_2,
+    distribution = qunif,
+    method = "ell",
+    alpha = .05,
+    points_params = list(col="red", type="l"),
+    log10 = TRUE
+  )
R> # Add the Q-Q plot for scenario 1.
R> lines(-log10(c(1:n) / (n + 1)), -log10(sort(pvals_scenario_1)), col = "blue")
R> legend(
+    "topleft",
+    legend = c("s=20","s=200"), 
+    col = c("red","blue"), 
+    lty = 1
+  )
\end{Sinput}
\end{Schunk}

@@

\subsection{Testing for Normality of Regression Residuals}
When performing an ordinary least squares regression (OLS), it is common to assume that the error terms are idependentally drawn from a normal distribution. Specifically, suppose that we have $n$ observations $\{\mathbf{x_{i}}, y_{i}\}_{i=1}^{n}$, where each $y_{i}$ is a scalar and each $\mathbf{x_{i}}$ is a $1xp$ vector of regressors. Then, we model each $y_{i}$ as
\begin{equation*}
  y_{i} = \mathbf{x_{i}}^{T}\boldsymbol\beta + \epsilon_{i}
\end{equation*}
where $\mathbf{\boldsymbol\beta}$ is a $1xp$ vector of regression parameters to be estimated and we assume
\begin{equation}
  \epsilon_{1}, ..., \epsilon_{n} \distras{iid} N(0, \sigma^{2}).
\end{equation}
Once $\boldsymbol\beta$ and $\sigma^{2}$ are estimated (typically via maximum liklihood), for each $y_{i}$ we define the residual as 
\begin{equation*}
  r_{i} = \mathbf{x_{i}}^{T}\mathbf{\hat{\boldsymbol\beta}} - y_{i},
\end{equation*}
where $\hat{\boldsymbol\beta}$ is our estimate of $\boldsymbol\beta$. Then, we would like a Q-Q plot of the residuals with a Normal reference distribution to aid in testing assumption (4) above. Strictly speaking, even if assumption (4) is correct, the residuals are not independent because OLS leads to an estimation of $\hat{\boldsymbol\beta}$ such that $\sum_{i=1}^{n}r_{i} = 0$. However, accounting for the resulting negative correlation between the residuals when creating testing bands would be very difficult, and with a sufficiently large sample size this correlation becomes negligible. Thus, without an a priori reason to believe that the errors may deviate from normality at any particular point in the distribution, it makes sense to use ELL bands in this case. This is very easy to do with \pkg{qqconf}, as we show below.
\newline
\newline
First, we generate data to perform a regression. Here, we will generate each $\epsilon_{i}$ independently from a $t(2.5)$ distribution. 
%
\begin{Schunk}
\begin{Sinput}
R> set.seed(100)
R> n <- 100
R> x <- runif(n)
R> eta <- rt(n, df = 2.5)
R> y <- x + eta
\end{Sinput}
\end{Schunk}
%
Then, we fit a regression with the simulated data
%
\begin{Schunk}
\begin{Sinput}
R> reg <- lm(y ~ x)
\end{Sinput}
\end{Schunk}
%
With \pkg{base}-\proglang{R} functionality, we can can create a qqplot using the function \code{qqnorm} and \code{qqline}. 
%
%
\begin{Schunk}
\begin{Sinput}
R> qqnorm(reg$residuals)
R> qqline(reg$residuals)
\end{Sinput}
\end{Schunk}
\includegraphics{manuscript-base_qqplot}
%
%
\newline
Clearly there's some indication of deviation in this plot, but it can be hard to tell how significant the deviation is without confidence bounds. With our function \code{qq_conf_plot}, it is much easier to be able to judge the significance of the deviation.
%
%
\begin{Schunk}
\begin{Sinput}
R> library(qqconf)
R> qq_conf_plot(
+    obs = reg$residuals,
+    points_params = list(col="red")
+  )
\end{Sinput}
\end{Schunk}
\includegraphics{manuscript-qqconf_qqplot}
%
%
\newline
Here, we can clearly see that in the right tail of the residuals goes beyond the normal testing bounds, giving strong evidence that the errors were not generated from a normal distribution.
